Grove: A Convergent Collaborative Structure-Editor Calculus

### 2020-11-06 grove-paper.tex
Edits and commutativity merge (Michael D. Adams)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% \figureCommutativity
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\figureCommutativity}{
\begin{figure}
\centering

@@

@@

TODO: explain red, green, blue

@@

@@

\subsubsection{Movement Actions}
\label{sub:Move}

TODO: in this section, discuss how we have to be careful to avoid creating cycles

A consequence of our design is that a move operation
does not have to delete code from one position and create a copy
in another position.
Instead, by deleting just the edge to the code to be moved and then
adding a new edge to that code somewhere else, the actual code is
moved, not just a copy.

For example, \autoref{fig:Move} Alice
moves \texttt{\hole{} * y} from the left child of \texttt{+} to its right child.
This corresponds to the following sequence of graph edits, shown in \autoref{fig:Move}:
(a) delete \Edge{\eidWrapPlus} leaving \Vertex{\vidSimpleTimes} temporarily orphaned and 
(b) add an edge (\Edge{\eidMoveTimes}) to \Vertex{\vidSimpleTimes} in the right child position of \Vertex{\vidWrapPlus}.

This move semantics gives us a richer structure than
when treating code as a list of lines.  We exploit
this in Section~REF:TODO in order merge edits that involve moving code.

Of course, if the user wants a genuine copy
and not just a move, that can be achieved
by creating new vertexes with fresh identifiers.
(TODO: figure for this?)

This is similar to the paste with formatting versus
paste without formatting.

If only one restore, then it does what you expect (basically a cut and paste)

If multiple replicas restore to different places, then you get some complexity:

@@

@@

\figureCommutativity

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% \figureCommutativity
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\figureCommutativity}{
\begin{figure}
\centering
\begin{tikzpicture}
\path ( 0cm, 0cm) node (a) {Original\\Version};
% \path (-1cm,-1cm) node (b) [alice] {Alice's Version};
% \path ( 1cm,-1cm) node (c) [bob] {Bob's Version};
% \path ( 0cm,-2cm) node (d) {Combined Version};
% \path [draw,->,alice] (a) -- node [pos=0.7,auto,swap] {Alice's Edits} (b);
% \path [draw,->,bob]   (a) -- node [pos=0.7,auto]      {Bob's Edits}   (c);
% \path [draw,->,merge] (b) -- node [pos=0.3,auto,swap] {Merge} (d);
\path [draw,->,merge] (c) -- node [pos=0.3,auto]      {Merge} (d);
\end{tikzpicture}

@@

TODO: move into section 2.3.1

TODO: star on multi-step arrows

@@

@@

TODO: explain cross-page arrows

@@

\subsubsection{Repositioning Actions}
\label{sub:Repositioning}

If a user want to move code from one place to another we have two options.
The first is to delete the code from its old position, and
reconstruct the code in its new position.
However, edges and vertices have unique identifiers in Grove 
to support the collaborative editing features we discuss next,
so reconstruction will not preserve identity of vertices and edges. 
Consequently, we include explicit repositioning actions that preserve identity.
These would be triggered by, for example, a drag and drop, or a cut and paste
(but not copy and paste, or a second paste after a cut, which do not need to
preserve identity and so would rely on reconstruction as already described).

In terms of the graph, a repositioning action involves simply deleting the edge
from the original parent then adding an edge from the new parent.
Thus the code is actually \textit{moved}, not copied.

For example, in \autoref{fig:Move} Alice continues her edits by
moving \texttt{\hole{} * y} from the left child of \texttt{+} to its right child.
This corresponds to the following sequence of graph edits, shown in \autoref{fig:Move}:
(a) delete \eWrapPlus leaving \vSimpleTimes temporarily orphaned and 
(b) add \eMoveTimes to \vSimpleTimes in the right child position of \vWrapPlus.

@@

TODO: (Place somewhere) This move semantics gives us a richer structure than
when treating code as a list of lines.  We exploit
this in Section~REF:TODO in order merge edits that involve moving code.

@@

%\figureCommutativity

\begin{figure}
\centering
\begin{tikzpicture}
\path ( 0cm, 0cm) node (a) [align=center]       {Original \\ Version};
\path (-1cm,-2cm) node (b) [align=center,alice] {Alice's  \\ Version};
\path ( 1cm,-2cm) node (c) [align=center,bob]   {Bob's    \\ Version};
\path ( 0cm,-4cm) node (d) [align=center]       {Combined \\ Version};
\path [draw,->,alice] (a) -- node [pos=0.7,auto,swap] {Alice's Edits} (b);
\path [draw,->,bob]   (a) -- node [pos=0.7,auto]      {Bob's Edits}   (c);
\path [draw,->,merge] (b) -- node [pos=0.3,auto,swap] {Merge} (d);
\path [draw,->,merge] (c) -- node [pos=0.3,auto]      {Merge} (d);
\end{tikzpicture}

TODO: move into section 2.3.1

TODO: star on multi-step arrows

\caption{TODO:Commutativity}
\label{fig:Commutativity}
\end{figure}
